
---
title: "Qualitative meta-analysis results"
author: "Amanida R package"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
params: 
  file_name: input_file
  separator: ";"
  analysis_type: "quan"
  column_id: coln
  pvalue_cutoff: 0.05
  fc_cutoff: 4
  votecount_lim: 1
  show_code: FALSE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra)
library(tidyverse)
library(amanida)
```

## Data inspection

Data is loaded using `amanida_read` function:

```{r, eval = F}
# Demo code with fixed parameters
coln = c("Compound Name", "P-value", "Fold-change", "N total", "References")
input_file <- system.file("extdata", "dataset2.csv", package = "amanida")
datafile <- amanida_read(input_file, mode = "quan", coln, separator=";")
```

Your data:

```{r, echo = F}
datafile <- amanida_read(params$file_name, params$analysis_type, params$column_id, params$separator) 
```

Here is shown 10 random rows:

```{r, echo = F}
data_imp <- datafile %>% 
  mutate(foldchange = round(foldchange,2),
         pvalue = format(pvalue, scientific = T, digits = 3),
         trend = case_when(trend == -1 ~ "Down-regulated",
                           T ~ "Up-regulated")) %>%
  rename("ID" = id,
         "P-value" = pvalue,
         "Fold-Change" = foldchange,
         "Number of participants" = N,
         "Reference" = ref,
         "Behaviour" = trend) %>%
  slice_sample(n = 10)

kbl(data_imp, 
      booktabs = T, 
      caption = "Table 1: Data exploration",
    align = "c") %>%
  kable_styling(latex_options = c("basic", "scale_down")) 
```

## Quantitative analysis

Quantitative analysis consist for each different identifier in combine the statistical significance and the effect size reported on the different studies. This combinations are made as follows:

* Statistical significance (p-value): Fisher's method weigthed by number of participants on the study.
* Effect size (fold-change): logarithmic transformation for average with weighting by number of participants. 

*Important: data is divided by trend before combination to respect discrepancies.*

`compute_amanida` function will be used to combine p-values and fold-change. 

```{r}
# Demo code with fixed parameters
amanida_result <- compute_amanida(datafile) 
```

Here are shown your results:

```{r, eval = F}
amanida_result@stat
```

```{r, echo = F}
kbl(amanida_result@stat %>%
      arrange(pval) %>%
      mutate(pval = format(pval, scientific = T, digits = 3),
             fc = round(fc, 2)) %>%
      rename("ID" = id,
             "Trend" = trend,
             "P-value combined" = pval,
             "Fold-change combined" = fc,
             "Total sample size" = N_total,
             "Reference" = reference), 
      booktabs = T, 
      caption = "Table 2: Quantitative analysis",
    align = "c") %>%
  kable_styling(latex_options = c("basic", "scale_down")) %>%
  scroll_box(width = "100%", height = "400px") 
```


### Significant results for quantitative analysis

```{r, echo = F}
sig_res <- amanida_result@stat %>%
  filter(pval < params$pvalue_cutoff,
         fc > params$fc_cutoff | fc < 1/params$fc_cutoff)
```
The cut-off for significant results are:

* p-value: `r params$pvalue_cutoff`.
* fold-change: `r params$fc_cutoff` for up-regulated and `r 1/params$fc_cutoff` for down-regulated.

Using indicated cut-offs there are **`r nrow(sig_res)`** compounds significant, `r length(which(sig_res$trend == 1))` up-regulated and `r length(which(sig_res$trend == -1))` down-regulated compounds.

### Graphical results

To see graphically the results obtained is used `volcano_plot`:

```{r, eval = F}
# Demo code with fixed parameters
volcano_plot(amanida_result, cutoff = c(0.05,4))
```

With `r params$file_name` and a p-value cut-off of `r params$pvalue_cutoff` and fold-change cut-off of `r params$fc_cutoff` (up-regulated) or `r 1/params$fc_cutoff` (down-regulated):

```{r, out.width= '100%', echo = F}
volcano_plot(amanida_result, cutoff = c(params$pvalue_cutoff, params$fc_cutoff))
```

*Note: labeled compounds are the ones found significant for the cut-off indicated*

As we have divided data by trend, now will be inspected the behaviour to see if there are discrepancies on the results obtained. This is evaluated comparing the number of articles where a element is found up-regulated and down-regulated. It is also ploted the total results, as vote-counting, obtained by the difference between up-regulated and down-regulated number of reports.  

As dataset is big, on the following plot will be only displayed elements bigger than `r params$votecount_lim` votes and the elements with discrepancies between studies:

```{r, eval = F}
# Demo code with fixed parameters
explore_plot(datafile, type = "mix", counts = 2)
```

Using counts = `r params$votecount_lim`:

```{r pressure, echo=FALSE}
explore_plot(datafile, type = "mix", params$votecount_lim)
```

It is important to check if our significant compounds have discrepancies on the reports. 

## Qualitative results

To see the complete results for vote-counting use:

```{r, eval = F}
amanida_result@vote 
```


```{r, echo = F}
kbl(amanida_result@vote %>%
      arrange(desc(abs(votes))) %>%
      mutate(vote_counting = round(vote_counting, 2)) %>%
      rename("ID" = id,
        "Votes" = votes,
             "Articles reporting" = articles,
             "Votes/articles" = vote_counting), 
      booktabs = T, 
      caption = "Table 3: Quanlitative analysis",
    align = "c") %>%
  kable_styling(latex_options = c("basic", "scale_down")) %>%
  scroll_box(width = "100%", height = "400px") 
```


Qualitative results can be also graphically inspected without trend division using:

```{r, eval = F}
# Demo code with fixed parameters
vote_plot(amanida_result, counts = 2)
```

Using counts = `r params$votecount_lim`:

```{r, out.width= '100%', echo = F}
vote_plot(amanida_result, counts = params$votecount_lim)
```


*Qualitative analysis using Amanida R package*



