
---
title: "Qualitative meta-analysis results"
author: "Amanida R package"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: lumen
params: 
  file_name: input_file
  separator: ";"
  analysis_type: "qual"
  column_id: coln
  votecount_lim: 1
  show_code: FALSE
---

```{r setup, echo=FALSE, warning = F, message = F}
set.seed(123)
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(kableExtra, include.only = c('kbl', 'kable_styling', 'scroll_box', 'footnote'))
library(amanida)
```

## Data exploration

Data is loaded using `amanida_read` function:

```{r, eval = F}
coln = c("Compound Name", "Behaviour", "References")
input_file <- system.file("extdata", "dataset2.csv", package = "amanida")
datafile <- amanida_read(input_file, mode = "qual", coln, separator=";")
```

Your data:

```{r, echo = F}
datafile <- amanida_read(params$file_name, params$analysis_type, params$column_id, params$separator) 
```

Here is shown the imported table:

```{r, echo = F}
data_imp <- datafile %>% 
  mutate(trend = case_when(trend == -1 ~ "Down-regulated",
                           T ~ "Up-regulated")) %>%
  rename("ID" = id,
         "Reference" = ref,
         "Behaviour" = trend)

kbl(data_imp, 
      booktabs = T, 
      caption = "Table 1: Data exploration",
    align = "c") %>%
  kable_styling(latex_options = c("basic", "scale_down")) %>%
  scroll_box(width = "100%", height = "400px")
```

## Qualitative analysis

Qualitative analysis is performed by the vote-counting computed as follows: for compounds up-regulated are count as +1 and for down-regulated as -1. Total sum is obtained for each compound. 

```{r, eval = T}
vote_results <- amanida_vote(datafile)
```

```{r, eval = F}
vote_results@vote
```

```{r, echo = F}
kbl(vote_results@vote %>%
      arrange(desc(abs(votes))) %>% 
      select("ID" = id, "Votes" = votes, "Articles reporting" = articles), 
      booktabs = T, 
      caption = "Table 3: Qualitative analysis",
    align = "c") %>%
  kable_styling(latex_options = c("basic", "scale_down")) %>%
  scroll_box(width = "100%", height = "400px") 
```

Qualitative results can be also graphically inspected without trend division. Here the vote_plot is set to show the identifiers with a minimum of a vote-counting of `r params$votecount_lim`

```{r, out.width= '100%', echo = T}
vote_plot(vote_results, counts = params$votecount_lim)
```

Here, the explore_plot will only display the relevant identifiers (elements bigger than 2 votes) or those with discrepancies between studies. 

As dataset is big, on the following plot will be only displayed elements bigger than `r params$votecount_lim` votes and the elements with discrepancies between studies:

```{r pressure, echo=T, message=TRUE}
explore_plot(datafile, type = "mix", params$votecount_lim)
```

*Qualitative analysis using Amanida R package*
